Topological Equivariant Artist Model for Visualization Library Architecture

### 2024-12-29 paper/main.tex
an artist is a natural transform on sheaves

Because Artists can be constructed as morphisms of sheaves over the same base spaces through the application of pushforward and pullback functors, they are natural transformations \cite{bradleyWhatNaturalTransformation}.
\begin{definition}
  Given two functors $F, G: \mathcal{C}\rightarrow \mathcal{D}$, a \textbf{natural transformation} $\alpha: F \rightarrow G$ is a function which assigns to each object $c$ of $\mathcal{C}$ a morphism $\alpha_c:F(input) \rightarrow G(c), G(c) \in \mathcal{D}$, in such a way that for every morphism $f:c \rightarrow c^\prime, c^\prime \in \mathcal{C}$, the morphisms in $\mathcal{D}$ commute such that $\alpha_c^{\prime}(F(f)(F(c))) = G(f)(\alpha_c(F(c))$. When this holds, $\alpha_{c}$ is \textit{natural} in $c$.\cite{maclaneCategoriesWorkingMathematician2013}.
\end{definition}.

@@

When the data and graphic containers are implemented as sheaves, the artist preserves the sheaf constraints because it is using the sheaf constraint satisfying map $\vindex$ to maintain the correspondence between data and graphing indexing spaces. Given a subset of the index $\dbase_2 \xhookrightarrow{i} \dbase_1$, all the data in the larger index is included in the smaller set $\Gamma(\dbase_1, \dtotal) \xrightarrow{i^*} \Gamma(\dbase_2, \dtotal)$ such that visualization specs of the smaller set must include specs for the larger set of data values  $\vartist(\Gamma(\dbase_1, \dtotal)) \xrightarrow{i^*} \vartist(\Gamma(\dbase_2, \dtotal))$. While the curves in \autoref{fig:atct:presheaf} are meant as stand-ins for the data, they also serve to illustrate that the relationship is identical in the graphic space such that there may be more graphics available for data from a smaller indexing space because there may be more values on the smaller indexing space. The artist also preserves the \textit{locality} axiom in that identical datasets $\dsection^{a} = \dsection^{b}$ generate identical specifications $\vartist(\dsection^a) = \vartist(\dsection^b)$. The artist also respects \textit{gluing} in that graphics are composable, as described in \autoref{sec:artist:operators}, such that a set of artists that generate graphics for parts of a dataset are equivalent to an artist that generates a graphic for the whole dataset. Managing subsets of data and on demand data in a structure preserving way is necessary to support visualizing streaming data \cite{krstajicVisualizationStreamingData2013} and techniques that rely on shifting or resampling the indexing space, such as  panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}.

>>>>

Because Artists can be constructed as morphisms of sheaves over the same base spaces through the application of pushforward and pullback functors, they are natural transformations \cite{bradleyWhatNaturalTransformation}.
\begin{definition}\label{def:natural-transform}
  Given two functors $F, G: \mathcal{C}\rightarrow \mathcal{D}$, a \textbf{natural transformation} $\alpha: F \rightarrow G$ is a function which assigns to each object $c$ of $\mathcal{C}$ a morphism $\alpha_c:F(input) \rightarrow G(c), G(c) \in \mathcal{D}$, in such a way that for every morphism $f:c \rightarrow c^\prime, c^\prime \in \mathcal{C}$, the morphisms in $\mathcal{D}$ commute such that $\alpha_c^{\prime}(F(f)(F(c))) = G(f)(\alpha_c(F(c))$. When this holds, $\alpha_{c}$ is \textit{natural} in $c$.\cite{maclaneCategoriesWorkingMathematician2013}.
\end{definition}.

@@

Specifying the artist as a natural transform (\autoref{df:natural-transform}) on sheaves means that the artist preserves the presheaf and sheaf constraints (\autoref{def:atct:presheaf}, \autoref{def:atct:sheaf}) when mapping data into graphics. The set of possible visualizations for a larger set of values must include all visualizations for a subset of those values  $\vartist(\Gamma(\dbase_1, \dtotal)) \xrightarrow{i^*} \vartist(\Gamma(\dbase_2, \dtotal))$ when more values may exist on the subset of the index $\dbase_2 \xhookrightarrow{i} \dbase_1$. While the curves in \autoref{fig:atct:presheaf} are meant as stand-ins for the data, they also serve to illustrate that the relationship is identical in the graphic space. The artist preserves the \textit{locality} axiom in that identical datasets $\dsection^{a} = \dsection^{b}$ generate identical specifications $\vartist(\dsection^a) = \vartist(\dsection^b)$. The artist respects \textit{gluing} in that graphics are composable, as described in \autoref{sec:artist:operators}, such that a set of artists that generate graphics for parts of a dataset are equivalent to an artist that generates a graphic for the whole dataset. Managing subsets of data and on demand data in a structure preserving way is necessary to support visualizing streaming data \cite{krstajicVisualizationStreamingData2013} and techniques that rely on shifting or resampling the indexing space, such as  panning and zooming \cite{NekrasovskiEvaluationPanZoom2006}.
